Formalizing Stack Safety as a Security Property

### 2023-01-17 submission5/core.tex
In the processing of rearranging the paper

\usepackage{xcolor,listings}

@@

The term {\em stack safety} is associated with a
variety of compiler, run-time, and hardware mechanisms for protecting stack
memory. But, unlike ``the heap,'' the stack does not correspond to a single
high-level language concept, and mechanisms for protecting it
typically lack precise specifications,
relying instead on informal descriptions and examples of bad
behaviors that they prevent. A more rigorous specification is especially desirable
for unusual enforcement mechanisms\bcp{why especially for this?}
like the ``lazy'' stack safety
micro-policies proposed by Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}.
% The term {\em Stack safety} is associated with a
% variety of compiler, run-time, and hardware mechanisms for protecting stack
% memory. But unlike heap memory, the stack does not correspond to any single
% high-level language concept. The result is that
% these mechanisms typically lack precise specifications,
% relying instead on informal descriptions and examples of the bad
% behaviors that they prevent. Such a specification is especially desirable
% for an unusual enforcement mechanism, like the ``lazy'' stack safety
% micro-policies proposed by Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}.

We propose a formal characterization of ``observable'' stack safety based on
concepts from language-based security. Stack safety is a
combination of an integrity property (``the private
state in each caller's stack frame is held invariant by the callee'')
and\bcp{two?} confidentiality properties\bcp{plural?} (``the caller's
and callee's behavior are insensitive to one another's private
state,''), both compatible with lazy tagging\bcp{undefined term; also,
the fact that we've mentioned tagged enforcement twice already makes
it really feel like we are proposing an abstract theory of exactly one
thing}, plus a control-flow property.

The stack governs a large number of language features that are often excluded
from discussion of stack safety. Our stack safety property targets a system with
both caller- and callee-saved registers, coroutines, arguments passed on the stack,
exceptions, [tail call elimination, heaps, and concurrency.]

@@

The protections offered by such mechanisms are commonly described by giving
concrete examples of attacks that they can prevent---corruption of return
addresses, buffer overflows, use of uninitialized variables, etc.---leaving
a more formal characterization to the reader's intuition. Previous attempts
to produce a formal characterization do so using a {\it fully abstract overlay semantics},
a style of correct-by-construction machine \cite{SkorstengaardSTKJFP}.
The latest stack-safety overlay semantics \cite{Georges22:TempsDesCerises} capture
both control-flow and data protection (confidentiality and integrity).
But, this approach is not suitable for {\em lazy} enforcement mechanisms like those
of Roessler and DeHon, in which stores are not checked, and illegal stores are
only caught when their location is later read. We aim to provide a stack-safety
property suitable for testing lazy enforcement.

@@

Stack safety is challenging to specify in full generality, because the stack is fundamentally
an architecture-specific construct. Its desired behavior varies based on ISA, calling convention,
and the features of whichever programming language originated the code that must be protected.
Our approach centers around a tagged architecture, like PIPE \cite{???} or STAR \cite{???},
particularly using a lazy micro-policy. However, our design is factored into three pieces:
(1) the definition of {\em security relevant operations} in a given machine,
(2) the definition of a {\em security context} which tracks the security needs of different
elements of the state over the course of execution, and
(3) the definition, for a given security context, of how we distinguish secure behavior
from insecure. This separation of concerns allows (3) to be relatively general, while
architecture-specifics are isolated in (1) and some special features require extending
(2).

A key technical novelty in these definitions,
compared to standard formulations of confidentiality and integrity from the
security literature, is that they
are ``nested'': {\em each} caller is guaranteed protection from its
immediate callees (which, in turn, are protected from their immediate
callees, etc.).
Caller confidentiality is especially interesting: It is based on a traditional
notion of noninterference, but, whereas ordinary noninterference
is an end-to-end hyper-property on whole program runs, caller
confidentiality is a nested form of noninterference that applies to
part of a run, requiring that the callee’s behavior be
invariant under hypothetical scrambling of the caller’s stack
but allowing the caller to access its own data upon return.

Our threat model is very strong: it allows the attacker
to execute arbitrary code, including attempting to smash
the stack to disrupt the program’s control flow. Our stack safety
properties promise that, even in the presence of such attackers, confidentiality
and integrity still apply.
%

For confidentiality and integrity, our properties\bcp{by this point,
  the reader may be getting confused about why we are talking about
  multiple properties, rather than just one (with multiple clauses)} are {\em observational},
meaning that they are only concerned with the observable behavior of the system.
A function activation might read from and write to other activations' data
without violating an observational property, as long as such reads and writes do not
affect the observable behavior of the
system. This permissiveness enables our properties to handle lazy
enforcement mechanisms.  \bcp{do we have any in mind aside from Nick's
tagging-based ones?}

To demonstrate the utility of our formal characterization, we use these
properties to validate and improve an existing enforcement mechanism, the
{\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
in the Coq proof assistant on top of a RISC-V specification.  We
use QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for Coq, to generate random programs and check
that Roessler and DeHon's micro-policies correctly detect the ones that
attempt to violate one of our properties. Furthermore, we
%
check that the testing framework is able to generate counterexamples
that violate our properties but are \emph{not} halted by incorrect
variants of the enforcement mechanisms, including ones that we accidentally created
while implementating the micro-policy and ones that we
intentionally crafted to increase our confidence in the effectiveness
of our testing.

We find that Roessler and DeHon's {\em Depth Isolation} micro-policy, in
which memory within each stack frame is tagged with the depth of
the function activation that owns the frame and access is
permitted only when an activation at that depth is currently executing, validates our
stepwise properties. On the other hand, our testing reveals that \emph{Lazy Tagging and Clearing}
violates the temporal aspect of confidentiality in
cases where data can leak across repeated calls to the same callee,
and also violates integrity if the leak uses the caller's frame. We
propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
confidentiality, albeit at some performance cost.

@@

\begin{itemize}
\item We formalize permissive stack confidentiality and integrity properties.
  Both are parameterized over a notion of external
  observation, and are violated only if accessing secrets or overwriting
  data causes a visible change in system behavior %(\cref{sec:lazy}).
\item We apply\bcp{instantiate?} these definitions to a complex\bcp{is
  its complexity a selling point?} system
  with features such as argument passing
  on the stack, callee-saves registers, exceptions, and a simple
  coroutine system% (\cref{sec:ext}).
  \bcp{Are these features really independent of the generic model?  If
  so, that's a nice point that should be brought out clearly.}
\item We use property-based random testing to validate the relationship between
  our properties and micro-policies. %(\cref{sec:testing})
  \bcp{Micro-policies have not been explained, and we haven't
    motivated why this would be an important or useful thing to do.}
\end{itemize}
%\Cref{sec:relwork,sec:future} discuss related and future work.
\bcp{Why did the section references get taken away from all that?  And
why delete the last sentence?}

Our artifact contains Coq formalizations of our properties for illustrative purposes
as well as the testing implementation.
\bcp{Make clear that there are no proofs!}

  We state our properties in terms of a generic machine, making minimal assumptions
  on the machine architecture and the nature of its built-in safety enforcement mechanisms
  (if any).  Our machine model is meant to be a faithful description of the\bcp{??} hardware,
  with a semantics that covers both ordinary instruction behavior and the behavior of
  any tags, capabilities, etc. that the machine might support.
  This semantics is intentionally \emph{not} ``safe-by-construction;'' it gives
  no guarantee that the enforcement hardware is being used correctly to achieve desired
  security properties.

  On top of this machine model, we layer a {\it security semantics}, which extends the
  underlying machine with additional context about the identity of current and pending
  functions (which act as security principals) and about the registers and memory they require
  to be secure. This added context is purely notional;
  it does not affect the real machine. The security context
  evolves dynamically through the execution of {\it security-relevant operations},
  which include events like calls, returns, and frame manipulations.
  Our security properties are phrased in terms of this context, often as predicates
  on future states, e.g. of the form ``when control returns to the current function...'',
  or relations on traces of future execution (hyper-properties).

  The security-relevant operations typically correspond to underling machine instructions,
  but the correspondence may not be obvious just by inspecting the machine code.
  For example, in the tagged RISC-V machine we use in our examples and tests,
  calls and returns are conventionally performed by {\tt jal} (``jump-and-link'')
  and {\tt jalr} (``jump-and-link-register'') instructions, respectively, but these
  instructions might also be used for other things. We therefore assume that there
  is some external mechanism for distinguishing which instructions are security-relevant;
  the most obvious technique is to add optional labels to instructions, and that is
  what we do in this paper, but the theory does not depend on this choice.

  Our security semantics has similarities to the overlay semantics that have been proposed
  to characterize stack safety in the past~\cite{SkorstengaardSTK}, but, like the basic machine
  semantics, it is not safe-by-construction. Rather, it gives us a means to \emph{define}
  the security properties that a correct enforcement scheme should implement, using
  some combination of hardware and software.

In the next section, we give a detailed example of a security semantics for a simple setting
in which our security-relevant operations are restricted to calls,
returns, and allocation of
private memory within the current function activation. In later sections, we extend this model
to the full system that we test, which features
% \begin{itemize}
% \item
function calls and returns, with caller- and callee-saved registers,
% \item
allocation of strictly private and strictly public regions on the stack,
% \item
arguments spilled onto the stack and/or passed by reference,
% \item
exceptions,
and
% \item
tail-call elimination.
% \end{itemize}

Finally, we give a property definition for stack safety in which (some) stack-allocated
objects are governed by a provenance-based capability system, accessible only to those
functions which have received a valid pointer to them.  \bcp{???}

%In section (TBD) we discuss the extension of this model into a simple concurrency setup.

@@

Our concurrency model is fairly
simplistic, assuming a fixed number of threads each with its own dedicated processor.
\sna{There is no concurrency in this document. Attempting to build it in from the beginning
  was rather complicated. The question is where to put that extra complexity? It's not difficult
  from a technical standpoint, subject to the one-thread-per-processor constraint.}\bcp{Appendix?}
We model memory-safe stack objects, but not a heap. Regions outside of
stacks can be used however the compiler likes, including as a heap, but no protection is
built in and our properties assume that if a pointer to a stack object is stored there,
it is permanently compromised.

Security-relevant operations consist of calls, returns, and allocations.
Calls identify the address of the call's target,
plus the argument registers. Allocations and deallocations are phrased in terms of a pair
of an offset from the stack pointer and a size. Returns don't require any additional information.
%
\begin{align*}
  \psi \in \Psi \ \ ::= \ \ & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} & \\
  | \ & \mathbf{return} & \\
  | \ & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
  | \ & \mathbf{dealloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N}
\end{align*}
%
The security principals in our account of stack safety are function
activations: at any given time, pending activations
have security requirements that are recorded in a call stack, and the current
activation tracks context information that will inform its requirements when
it makes a call. Each activation has a {\it view}
of the system that maps each state element to a {\it security class}:
it may be private to an inactive principal (\(\sealed\)),
in an allocated object that is accessible to the current principal (\(\object\)),
available to be allocated (\(\unsealed\)), or
always accessible (\(\public\)). The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
all other locations to \(\public\), and registers based on which set they
belong to: % (Figure \ref{fig:RISCVregs}):
\(\sealed\) for callee-saved, \(\unsealed\) for caller-saved, and \(\public\) otherwise.
%Elements of \(\mathcal{P}\),
%such as the program counter, are always \(\public\). Even if they might deserve their
%own protection, that protection is orthogonal to stack safety.

A {\it context} is a pair of the current activation's view and
a list of views representing the call stack (pending inactive
principals), ranged over by \(\sigma\).
%\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ VIEW}\]
The initial context is \(\context_0 = (V_0, \emplist)\).

As the machine takes a step that executes a particular operation, it updates the
corresponding context, as described informally in the following section, and formally
in Section \ref{sec:props}.

%\section{Calls, Returns, and Private Allocations}
\section{Properties by Example}

In this section we consider a simple setting with calls, returns, and private allocations,
on a RISC-V machine with a PIPE tagging system\bcp{undefined}. Here
\(\mathcal{V} ::= \mathcal{W} \times \mathcal{T}\),
for some type \(\mathcal{T}\) of tags to be elaborated later, with \(|(w,t)| = w\).
Figure \ref{fig:RISCVregs} gives an example division of the RISC-V registers into
security classes.
Figure \ref{fig:main} gives the C code and corresponding compiled RISC-V code
for a function {\tt main}. Code is compiled using the standard RISC-V
calling conventions~\cite{??}; the stack grows towards lower addresses. We assume that
{\tt main} begins at address 0 and its callee {\tt f} at address 100.

@@

While it would be simple to define integrity as ``all sealed elements retain their
values after the call,'' this would be stricter than necessary for many purposes.
For example, it arguably doesn't matter if {\tt f} overwrites {\tt main}'s data provided
that {\tt main} never reads that corrupted data.
To illustrate why this matters, let's consider two integrity
micro-policies: ``Depth Isolation'' and
``Lazy Tagging and Clearing.''

Depth Isolation tracks the depth of the current activation via a tag on the program counter.
Stack frames are initialized and cleared by writing
to every address in them, so that during a call at depth \(d\), the current activation's
frame is tagged with \(d\). Loads and stores targeting addresses in frames failstop
if the program counter tag does not match the frame tag. In this setting, at depth \(d\),
all memory tagged with a depth less than \(d\) corresponds to \(\sealed\) memory,
and should be indeed be inaccessible.

On the other hand, Lazy Tagging and Clearing skips the (inefficient)
initialization and clearing steps, and only updates tags when memory is written to.
It allows a callee to overwrite sealed memory, but that write tags the corrupted memory
with the callee's identity. Only if the corrupted memory's owner (or any other activation) tries
to read it will the micro-policy failstop.

Lazy Tagging and Clearing illustrates the value of ``extensional'' properties.
It is common to specify the behavior of a system, including security properties, in
terms of traces of observable events rather than details of the machine state.
Lazy enforcement works because the deferred check catches errors before they appear
in the trace. The nature of traces varies based on the properties of interest, but
given a particular model of observation, we can consider
corruption by {\tt f} acceptable as long as it does not affect {\tt main}'s observable behavior.

@@

\sna{Version that is less tied to lazy enforcement. It's still rather tied, actually,
  more than I remembered.}

One or more elements of a given state are {\it safe} if their values
in that state have no observable influence on future execution. We define this
formally as a sort of non-interference: for a state \(\sigma\) and
a set of elements \(\components\), \(\mathit{safe} ~ \sigma ~ \components\)
holds if we can replace the contents of \(\components\) with arbitrary values
and have the resulting state produce an equivalent trace. The state created
by modifying \(\components\) in this way is termed a variant.}

@@

In our example setting, the observation trace consists of the sequence
of values written to {\tt out}.\bcp{Is this the first place we've
  discussed what the observations actually are?  Is this the only
  thing they could be, or does the model support other options?}
The example in Figure \ref{subfig:integrity} modifies the value of {\tt sensitive},
which is \(\sealed\). Figure \ref{fig:variant} shows the state just after the call at step 5,
assuming that {\tt sec} is 5. Similar to WBCF, we will define
\(\intProp\) as a predicate on states that holds if
all relevant sealed addresses in \(\mach\) are the same as after step 5.
We once again require that \(\intProp\) hold on the state following the matching return,
which is reached by step 9. Here {\tt sensitive} has obviously changed, but is it relevant?
Consider a variant state in which {\tt sensitive} has any other value, arbitrarily
choosing 43 (and keeping its tag, \(t_2'\), unchanged.) As execution continues after the return
from the original state, it passes the {\tt if}-test on {\tt sensitive}, whereas the execution
from the variant does not, resulting in differing outputs. Therefore {\tt sensitive} \emph{is} relevant,
so \(\intProp\) does not hold, and integrity has indeed been violated.

@@

@@

\subsection{Facts About Calls and Returns}
\label{sec:facts}

@@

\begin{align*}
  \psi \in \Psi ::= & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa} &
  sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
  | & \mathbf{tailcall} ~ \addr_{target}  ~ \overline{\reg_{args}} ~ \overline{sa} &
  sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
  | & \mathbf{return} & \\
  | & \mathbf{alloc} ~ \mathit{pub} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N}, \mathit{pub} \in \mathbb{B} \\
  | & \mathbf{dealloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
\end{align*}

@@

\begin{align*}
  \psi \in \Psi ::= & \dots \\
  & \mathbf{promote} ~ \reg ~ sa & sa \in \REGS \times \mathbb{Z} \times \mathbb{N} \\
  & \mathbf{propagate} ~ \component_{src} ~ \component_{dst} \\
  & \mathbf{clear} ~ \component \\
\end{align*}

@@

%\section{Enforcement}
%\label{sec:enforcement}

%We use our properties to test the stack safety micro-policies of Roessler and DeHon,
%particularly their Lazy Tagging and Clearing. Our underlying machine is a RISC-V
%based on the MIT RISC-V formalism \cite{} and extended with {\it metadata tags},
%represented as additional state in \(\mathcal{P}\), one for each register, address,
%and the program counter:
%\[\mathcal{P} ::= t_\reg | t_\addr | t_\PCname\]
%
%We divide values into normal memory words and tags drawn from a policy-specific
%set of tags \(\mathcal{T}\), \(\mathcal{V} ::= \mathcal{W} | \mathcal{T}\).
%The machine's step function corresponds to that of the basic RISC-V machine,
%with the behavior of the tags at each step parameterized by a {\it micro-policy},
%which may turn a normal step into an error (represented by returning to the same state.)

%\paragraph*{Protecting Enforcement State}

%State beyond registers and memory is always considered \(\public\), and in a PIPE machine this
%includes tags. That does not mean that functions should be modifying them! It just means that we
%do not consider them to be part of the state being protected. It is less obvious that a function
%should be prevented reading a tag, and our properties do not rule it out, although in practice
%PIPE does not allow such reads. It would be quite possible to implement
%the properties so that they would forbid these activities, for instance by structuring \(\mathcal{P}\)
%to treat tags separately and apply stack safety to the tags on the stack, but there is no obvious
%application for such a stringent property---the enforcement state must be protected in order that
%enforcement not be subverted, not for its own sake.

\section*{[Meta] Remainder}

These document sections lay out a way to present our stack safety properties to build intuition and
then introduce the formalism. Next the full paper will need the testing section and an expanded
enforcement section, but these can largely be borrowed from previous submissions.

>>>>

\usepackage[table]{xcolor}

\usepackage{listings}

@@

The term {\em stack safety} is associated with a
variety of compiler, run-time, and hardware mechanisms for protecting stack
memory. But, unlike ``the heap,'' the stack does not correspond to a single
high-level language concept, making a formal specification elusive.
Protection mechanisms typically lack precise specifications, relying instead
on informal descriptions and examples of bad behaviors that they prevent.

When the prevented behaviors are straightforward, describing them may be sufficient
to justify an enforcement mechanism's effectiveness. Less intuitive enforcement
mechanisms instead leave us to ask: what security do we actually hope to gain,
and are we successful?

In this paper we focus on one such non-intuitive enforcement mechanism,
the ``lazy'' stack safety micro-policies proposed by
Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}. These permit functions
to write into one another's frames, but taint the result so that the frame's
owner cannot access it. No existent characterization of stack safety
captures this style of safety.

We propose a formal characterization of ``observable'' stack safety based on
concepts from language-based security. Stack safety is a
combination of an integrity property (``the private
state in each caller's stack frame is held invariant by the callee''),
two confidentiality properties (``the caller's
and callee's behavior are insensitive to one another's private
state,''), and a control flow property. By defining the properties
in terms of the observable behavior of the system, we expand the
space of supported enforcement mechanisms to include those that
allow certain (intuitively incorrect) behaviors but prevent them
from causing harm.

The stack governs a large number of language features that are often excluded
from discussion of stack safety. Our properties support a system with
both caller- and callee-saved registers, arguments passed on the stack,
exceptions, [tail call elimination, and concurrency.] They are modular by
design, in order to be further extensible to other features and architectures.

@@

The protections offered by such mechanisms are commonly described by giving
concrete examples of attacks that they can prevent---corruption of return
addresses, buffer overflows, use of uninitialized variables, etc.---leaving
a more formal characterization to the reader's intuition. Previous attempts
to produce a formal characterization do so using a {\it fully abstract overlay semantics},
a style of correct-by-construction machine \cite{SkorstengaardSTKJFP}.
But this approach is not suitable for the {\em lazy} enforcement mechanism
of Roessler and DeHon, in which stores are not checked, and illegal stores are
only caught when their location is later read. We aim to provide a stack-safety
property suitable for testing lazy enforcement.

@@

To demonstrate the utility of our formal characterization, we use these
properties to validate and improve an existing enforcement mechanism, the
{\em stack safety micro-policies} of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, re-implemented
in the Coq proof assistant on top of a RISC-V specification.  We
use QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for Coq, to generate random programs and check
that Roessler and DeHon's micro-policies correctly detect the ones that
attempt to violate one of our properties. Furthermore, we
check that the testing framework is able to generate counterexamples
that violate our properties but are \emph{not} halted by incorrect
variants of the enforcement mechanisms, including ones that we accidentally created
while implementating the micro-policy and ones that we
intentionally crafted to increase our confidence in the effectiveness
of our testing.

We find that Roessler and DeHon's \emph{Lazy Tagging and Clearing}
violates the temporal aspect of confidentiality in
cases where data can leak across repeated calls to the same callee,
and also violates integrity if the leak uses the caller's frame. We
propose a variant of {\em Lazy Tagging and Clearing} that testably enforces
confidentiality, albeit at some performance cost.

\subsection{Motivation}

Security mechanisms designed to prevent a certain class of attacks can be brittle---successfully
eliminating the attack as it exists at the time of publication, but leaving room for attackers
to find new, similar attacks. To mitigate this risk, mechanisms should aim to enforce formal
properties that can be proven or tested. Such properties become the specification against which
enforcement can be validated; even enforcement mechanisms that do not fulfill the property
benefit from the ability to articulate {\it why}.

Stack safety is challenging to specify in full generality, because the stack is fundamentally
an architecture-specific construct. (Contrast Acevedo de Amorim et al.'s \cite{} work on heap safety,
which is broadly consistent across architectures and therefore amenable to a very general approach.)
The desired behavior of the stack varies based on ISA, calling convention,
and the features of whichever programming language originated the code.
For instance, the current state-of-the-art stack-safety specification from Georges et al.
\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
control flow, and therefore defines {\it Well-bracketed Control Flow} (WBCF) under the assumption that
a callee should always return to its caller. But, in the presence of tail-call elimination, a
return from a callee to a distant ancestor is perfectly reasonable. Subtle tweaks to the definition
accompany any addition to the model.

We are not attempting to create a universal definition of stack safety for all systems.
Rather, we define the stack safety of a single, fairly realistic system, decomposed into
(1) a lightweight model of the system and its security-relevant features, from which we
derive the integrity, confidentiality, and control-flow requirements of each call and
return, (2) the application of that model to a particular target system, and (3)
the formal criteria for integrity, confidentiality, and control flow
to obey the requirements.

We treat stack safety as multiple properties, rather than one multi-faceted property, because
not all enforcement mechanisms attempt to implement all of them. Some focus exclusively on
control flow, others protect the confidentiality of the caller but not the callee, etc.

\paragraph*{Permissiveness, Laziness, and Tagged Architectures}

Security properties tend to be conservative, identifying violations that would be
harmless in practice. But enforcement mechanisms make trade-offs in the name of performance
and of compatibility with existing codebases. In the motivating case for this work,
Roessler and DeHon \cite{} present a collection of tag-based enforcement mechanisms
using PIPE \cite{???}, a tagged hardware reference monitor. In monitors such as PIPE and
STAR \cite{???}, every value in memory and registers is paired with a {\em metadata tag}.
A set of software-defined rules, termed a micro-policy, determines the tags for operation
results based on the tags of their inputs---or, if the input tags are in a dangerous
combination, it may halt the machine (a ``failstop.'')

Roessler and DeHon find that the obvious way to enforce stack safety in this setting
is inefficient when allocating and clearing stack frames, because each word must be initialized
with a tag corresponding to the (depth of the) current function activation. They propose
{\em Lazy Tagging and Clearing}, in which tags are not initialized at allocation, but only
when a function writes to memory. This means that such writes cannot be protected. Instead
the micro-policy enforces that memory be read by the same fuction that wrote it. This
lazy enforcement is incompatible with existing models of stack safety, particularly the
integrity side, because an attacker can modify another function's data. But the owner of
the data will never see the changes, so it may still be secure in some sense. We aim to elaborate
that sense, and to test whether the micro-policy fulfills it.
This key motivation leads us to present integrity and confidentiality as predicates on (pairs of)
machine states that make claims about the future observable behavior of execution from those states.

\subsection{Contributions}

@@

\begin{itemize}
\item We formalize permissive stack confidentiality and integrity properties.
  Both are parameterized over a notion of external
  observation, and are violated only if accessing secrets or overwriting
  data causes a visible change in system behavior.
\item We instantiate these definitions with a realistic
  system with features such as argument passing
  on the stack, callee-saves registers, and tail-call elimination.
  Our model is modular enough that adding these features is straightforward.
\item Our formalization of well-bracketed control flow in the presence of tail-call
  elimination is a slight novelty over existing models.
\item We use property-based random testing to validate the relationship between
  our properties and micro-policies.
\end{itemize}

Our artifact contains Coq formalizations of our properties for illustrative purposes
as well as their testing implementation. There are no proofs---we use primarily Coq for the
QuickChick library and to ensure that the formalizations are unambiguous.

\section{Key Ideas}

The paper is structured as follows. In \cref{sec:example}, we will walk through a function
call in a simple example machine and discuss how each of our properties applies to it,
informally. Then in \cref{sec:machine} we will introduce the formal machine model and give
its {\em security semantics}, described below. In \cref{sec:props} we give abbreviated
formal statements of the properties, while the full formalism is found in \cref{app:props}.
\Cref{sec:ext} gives the security semantics for the full, tested machine, and
\cref{sec:testing} describes the testing framework used to test it.
\Cref{sec:relwork,sec:future} discuss related and future work.
The remainder of this section introduces ideas that will be important throughout.

\paragraph*{Security Semantics}

A {\it security semantics} extends a machine
with additional context about the identity of current and pending
functions (which act as security principals) and about the registers and memory they require
to be secure. This added context is purely notional;
it does not affect the real machine. The security context
evolves dynamically through the execution of {\it security-relevant operations},
which include events like calls, returns, and frame manipulations.
Our security properties are phrased in terms of this context, often as predicates
on future states, e.g. of the form ``when control returns to the current function...'',
or relations on traces of future execution (hyper-properties).

The security-relevant operations typically correspond to underling machine instructions,
but the correspondence may not be obvious just by inspecting the machine code.
For example, in the tagged RISC-V machine we use in our examples and tests,
calls and returns are conventionally performed by {\tt jal} (``jump-and-link'')
and {\tt jalr} (``jump-and-link-register'') instructions, respectively, but these
instructions might also be used for other things. An instantiation of an operation
might require some external mechanism for distinguishing which instructions are
security-relevant; the most obvious technique is to add optional labels to instructions,
but the theory does not depend on this choice.

The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
the following table. Those that we explore in detail in Section \ref{sec:examples}
are shown in white: calls, returns, and the allocation and deallocation of local memory.
In the full formalism (Appendix \ref{app:full}) and the machine that we test,
we cover additional operations add parameters to our existing ones, to support
various forms of memory sharing and tail-call elimination. These are shown in light gray.
We give a proof-of-concept formalization of a more sophisticated, heap-like sharing model
based on pointer provenance in Appendix \ref{app:ptr}, represented by the dark grey
operations, but we do not test this.

\newcommand{\example}{\rowcolor{black!0}}
\newcommand{\testing}{\rowcolor{black!10}}
\newcommand{\theory}{\rowcolor{black!25}}

\begin{tabular}{| l | l |}
  \hline
  \(\psi \in \Psi\) & Parameters \\
  \hline
  \example \(\mathbf{call}\) & Target address \\
  \example & Argument registers \\
  \testing & Stack argument offset \& size \\
  \example \(\mathbf{return}\) & \\
  \example \(\mathbf{alloc}\) & Offset \& size \\
  \testing & Public flag \\
  \example \(\mathbf{dealloc}\) & Offset \& size \\
  \testing \(\mathbf{tailcall}\) & As \(\mathbf{call}\) \\
  \hline
  \multicolumn{2}{|c|}{{\it Pointer provenance operations}} \\
  \hline
  \theory \(\mathbf{promote}\) & Register, offset \& size \\
  \theory \(\mathbf{propagate}\) & Source register/address \\
  \theory & Destination register/address \\
  \theory \(\mathbf{clear}\) & Target register/address \\
  \hline
\end{tabular}

\vspace{\belowdisplayskip}

Our security semantics has similarities to the overlay semantics that have been proposed
to characterize stack safety in the past~\cite{SkorstengaardSTK}, but it does not restrict
the behavior of the underlying machine in any way. Rather, it tracks additional context
about the history of security-relevant operations, which inform the criterion a machine
must satisfy in order to correctly obey the properties.

\paragraph*{Security Classes}

The security principals in our account of stack safety are function
activations: at any given time, pending activations
have security requirements that are recorded in a call stack, and the current
activation tracks context information that will inform its requirements when
it makes a call. Each activation has a {\it view}
of the system that maps each state element to a {\it security class}:
it may be private to an inactive principal (\(\sealed\)),
in an allocated object that is accessible to the current principal (\(\object\)),
available to be allocated (\(\unsealed\)), or
always accessible (\(\public\)). The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
all other locations to \(\public\), and registers based on which set they
belong to: \(\sealed\) for callee-saved, \(\unsealed\) for caller-saved, and \(\public\) otherwise.

A {\it context} is a pair of the current activation's view and
a list of views representing the call stack (pending inactive
principals), ranged over by \(\sigma\).

\paragraph*{Variant States}

\paragraph*{Events and Traces}

It is common to specify the behavior of a system, including security properties, in
terms of traces of observable events rather than details of the machine state.
Lazy enforcement works because the deferred check catches errors before they appear
in the trace. The nature of traces varies based on the properties of interest, but
given a particular model of observation, we can consider
corruption by {\tt f} acceptable as long as it does not affect {\tt main}'s observable behavior.

@@

%Our concurrency model is fairly
%simplistic, assuming a fixed number of threads each with its own dedicated processor.
We model memory-safe stack objects, but not a heap. Regions outside of
stacks can be used however the compiler likes, including as a heap, but no protection is
built in and our properties assume that if a pointer to a stack object is stored there,
it is permanently compromised.

%\section{Calls, Returns, and Private Allocations}
\section{Properties by Example}
\label{sec:example}

In this section we consider a simple setting with calls, returns, and private allocations,
on a RISC-V machine with a PIPE tagging system. In this system a value is a pair of a
word (the ``payload'') and a tag drawn from some set \(\mathcal{T}\).
%\(\mathcal{V} ::= \mathcal{W} \times \mathcal{T}\),
%for some type  of tags to be elaborated later, with \(|(w,t)| = w\).
Figure \ref{fig:RISCVregs} gives an example division of the RISC-V registers into
security classes.
Figure \ref{fig:main} gives the C code and corresponding compiled RISC-V code
for a function {\tt main}. Code is compiled using the standard RISC-V
calling conventions~\cite{??}; the stack grows towards lower addresses. We assume that
{\tt main} begins at address 0 and its callee {\tt f} at address 100.

@@

While it would be simple to define integrity as ``all sealed elements retain their
values after the call,'' this would be stricter than necessary. Suppose that
a callee overwrites some data of its caller, but the caller never accesses that data.
By sheer luck, the callee has managed not to impact the observable execution beyond that one write.
Luck, or enforcement: Lazy Tagging and Clearing, from Roessler and Dehon, works by
failstopping any time a caller accesses callee-tainted data, so that out-of-frame
writes should never impact execution.

@@

@@

In our example setting, the observation trace consists of the sequence
of values written to {\tt out}.
The example in Figure \ref{subfig:integrity} modifies the value of {\tt sensitive},
which is \(\sealed\). Figure \ref{fig:variant} shows the state just after the call at step 5,
assuming that {\tt sec} is 5. Similar to WBCF, we will define
\(\intProp\) as a predicate on states that holds if
all relevant sealed addresses in \(\mach\) are the same as after step 5.
We once again require that \(\intProp\) hold on the state following the matching return,
which is reached by step 9. Here {\tt sensitive} has obviously changed, but is it relevant?
Consider a variant state in which {\tt sensitive} has any other value, arbitrarily
choosing 43 (and keeping its tag, \(t_2'\), unchanged.) As execution continues after the return
from the original state, it passes the {\tt if}-test on {\tt sensitive}, whereas the execution
from the variant does not, resulting in differing outputs. Therefore {\tt sensitive} \emph{is} relevant,
so \(\intProp\) does not hold, and integrity has indeed been violated.

@@

Security-relevant operations consist of calls, returns, and allocations.
Calls identify the address of the call's target,
plus the argument registers. Allocations and deallocations are phrased in terms of a pair
of an offset from the stack pointer and a size. Returns don't require any additional information.
%

%\[\context \in \CONTEXTS ::= \mathit{VIEW \times list ~ VIEW}\]
The initial context is \(\context_0 = (V_0, \emplist)\).

@@

@@

@@

@@

\section{Validation through Random Testing}
\label{sec:testing}

There are several ways to evaluate whether an enforcement mechanism enforces
stack safety properties. Ideally such validation would be done through formal proof over
the semantics of the enforcement-augmented machine.
However, while there are no fundamental barriers to producing such a proof,
it would be considerable work to carry out for a full ISA like RISC-V and
complex enforcement mechanism like the Depth Isolation micro-policy.
We therefore choose to validate the micro-policy of \cref{sec:enforcement} by
systematically \emph{testing} that it satisfies our properties.
This focus on testing is better aligned with our immediate
goal of evaluating real enforcement mechanisms for real machine architectures.
Formal proof remains future work.

We validate the micro-policy of the previous section by
systematically testing that it satisfies eager stack integrity and
confidentiality. We use a Coq specification of the RISC-V
architecture~\cite{Bourgeat2021AMF},
and extend it with a runtime monitor implementing a stack safety
micro-policy. We chose the Coq proof assistant as the setting for our implementation
to ensure that our coinductive trace definitions are well-formed, to reason about
them, and to leverage the power of the QuickChick property-based testing framework~\cite{Pierce:SF4}.

To use QuickChick, we build random test-case generators that produce
% \begin{enumerate}
% \item
(1) an initial RISC-V machine state, including most notably
  the program to be executed;
  % \item
(2)
  an initial policy state, tagging instructions corresponding to
  blessed call or return sequences appropriately, while marking
  all potential stack locations as $\tagNoDepth$; and
% \item
(3)
 a code map identifying the locations of calls.
% \end{enumerate}
